Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-08-06 LaTex/paper.tex
first draft for example section

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the negative binomial probability distribution (NBD) is well routed in theoretical arguments to model customer demand. Its parameters can be modelled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as SKU, opening day, and store location) and may depend on a multitude of describing variables or features such as sales location, weather, price, and so forth. It should be noted that these parameters are not independent between products. For example, a promotion applied to one product can, at least in principle, affect the sales of related products within the assortment. This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modelled holistically. Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance. The features may or may not differ between the mean and variance estimation models. It can also be beneficial to include the corresponding predicted mean from the mean model as feature in the variance model. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

@@

\subsection{Quantitative Evaluation of PDF Predictions}

@@

We use data from a Kaggle online competition about estimating unit sales of Walmart retail goods \cite{kaggle_data} to demonstrate forecasting of demand for different product-location-date combinations in form of full individual PDFs by means of Cyclic Boosting and subsequent qualitative and quantitative evaluation of the PDF predictions. The fields identifying a unique sample of the data set are store (\texttt{store\_id}) and item identifiers (\texttt{item\_id}) as well as date, while the field sales represents the target to predict. In the following, we use data from 2013-01-01 to 2016-05-22 for all 10 available stores and 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3}. Besides the key and target fields (\texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}), we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} to build features for our machine learning models.

@@

For our Cyclic Boosting mean estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several variables describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), and a flag denoting a promotion. We also include various two-dimensional combinations of the mentioned features, where one of the two dimensions being \texttt{store\_id} or \texttt{item\_id} to learn characteristics of individual locations and products. Furthermore, we include an exponential price-demand elasticity sub-estimator based on the ratio of reduced and normal price and individualized by \texttt{store\_id}. We plan to report on the inner workings of this method in a future work.

\textcolor{red}{... training and prediction periods ... residual correction ...}

\textcolor{red}{... results: mad, mean(y) ... example time series plot ...}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/mean_prediction}
\caption{\label{fig:mean_prediction} \textcolor{red}{time series of mean prediction and actuals for one item in all stores}}
\end{center}
\end{figure}

@@

For our Cyclic Boosting variance estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, a flag denoting a promotion, and the insample prediction of the mean model, and the individual correction factor of the residual correction step described in \ref{sec:example_mean}.

\textcolor{red}{... same training and prediction periods ... no residual correction ...}

\textcolor{red}{... pdf/cdf example for visualization...}

\begin{figure}
\begin{center}
\includegraphics[width=4cm]{../figures/pdf}
\includegraphics[width=4cm]{../figures/cdf}
\caption{\label{fig:pdf_example} \textcolor{red}{PDF/CDF example}}
\end{center}
\end{figure}

@@

\textcolor{red}{...}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/cdf_truth}
\caption{\label{fig:cdf_demand} \textcolor{red}{CDF histogram}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_yhat_mean}
\caption{\label{fig:invquant_mean} \textcolor{red}{invquant profile for mean predictions of all product-location-day combinations on x-axis}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_dayofweek}
\caption{\label{fig:invquant_dayofweek} \textcolor{red}{invquant profile for day of week of one product over all locations and days on x-axis}}
\end{center}
\end{figure}

\textcolor{red}{... quantitative CDF accuracy ...}

>>>>

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the negative binomial probability distribution (NBD) is well routed in theoretical arguments to model customer demand. Its parameters can be modelled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as SKU, opening day, and store location) and may depend on a multitude of describing variables or features such as sales location, weather, price, and so forth. It should be noted that these parameters are not independent between products. For example, a promotion applied to one product can, at least in principle, affect the sales of related products within the assortment. This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modelled holistically. Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance. The features may or may not differ between the mean and variance estimation models. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

@@

\subsection{Quantitative Evaluation of PDF Predictions}
\label{sec:cdf_acc}

@@

We use data from a Kaggle online competition about estimating unit sales of Walmart retail goods \cite{kaggle_data} to demonstrate forecasting of demand for different product-location-date combinations in form of full individual PDFs by means of Cyclic Boosting and subsequent qualitative and quantitative evaluation of the PDF predictions. The fields identifying a unique sample of the data set are store (\texttt{store\_id}) and item identifiers (\texttt{item\_id}) as well as date, while the field sales represents the target to predict. In the following, we use data from 2013-01-01 to 2016-05-22 for all 10 available stores and 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3}, where we use all data before 2016 for the training and all data from 2016 for the testing of our model. Besides the key and target fields (\texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}), we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} to build features for our machine learning models.

@@

For our Cyclic Boosting mean estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several variables describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), and a flag denoting a promotion. We also include various two-dimensional combinations of the mentioned features, where one of the two dimensions being \texttt{store\_id} or \texttt{item\_id} to learn characteristics of individual locations and products. Furthermore, to capture the price sensitivity as feature, we include an exponential price-demand elasticity sub-estimator based on the ratio of reduced and normal price and individualized by \texttt{store\_id}. We plan to report on the inner workings of this method in a future work.

Unlike most state-of-the-art time series forecasting methods, we do not include lagged target information, for example via stacking of exponentially weighted moving average features, in our model. The reason for this is that including such information makes the learning of exogenous effects, e.g. promotions or events, much harder, as the model tends to rely mainly on temporal confounding. So, to instead strive for direct learning of causal effects and in turn improve on explainability, mid- and long-term forecasting quality, and predictability of rare events, it is beneficial to not include lagged target information in the machine learning model. In order to capture recent trends that are not reflected in the exogenous features of the model, we apply an individual residual correction on each of the predictions of the machine learning model, which accounts for deviations between the exponentially weighted moving average of the predictions and targets of each product-location combination over the corresponding past. Again, we plan to give more insights on the reasoning for such a residual correction in a future work.

Just to give a rough verification for the accuracy of the mean predictions, we report the MAD and MSE averaged over the full used data set in the test period to be $1.68$ and $10.20$, respectively, while the average of the target, i.e. the sales, is $3.28$. \fig \ref{fig:mean_prediction} shows an example for the time series of both predictions and sales for one specific item, namely \texttt{FOODS\_3\_586} (which we will use throughout the rest of this example), summed over all 10 stores during the test period.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/mean_prediction}
\caption{\label{fig:mean_prediction} Time series of mean prediction and sales in test period for one specific item summed over all stores.}
\end{center}
\end{figure}

@@

For our Cyclic Boosting variance estimation model, the training and test periods are the same as for the mean estimation model, and the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}, meaning that the mean predictions are created in an insample way for the training period. We use a different set of features, namely the following: categorical variables for \texttt{store\_id} and \texttt{item\_id} and a flag denoting a promotion. An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_586} in store \texttt{WI\_1} on 2016-05-22 is shown in \fig \ref{fig:pdf_example}.

\begin{figure}
\begin{center}
\includegraphics[width=4cm]{../figures/pdf}
\includegraphics[width=4cm]{../figures/cdf}
\caption{\label{fig:pdf_example} Predicted PDF (left) and CDF (right) for a specific product-location-date combination.}
\end{center}
\end{figure}

@@

\fig \ref{fig:cdf_demand} shows the histogram of CDF observations according to the method described in sec. \ref{sec:cdf_histo} for item \texttt{FOODS\_3\_586} in the test period, where a significant deviation from the uniform distribution, expected for optimal PDF predictions, can be seen. The shape of the deviation points towards a slight underprediction of the mean estimation (see \fig \ref{fig:cdf_histos}), which is also visible in the time series plot of \fig \ref{fig:mean_prediction} (e.g. the high peak during two days end of April).

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/cdf_truth}
\caption{\label{fig:cdf_demand} Histogram of all observed CDF values of the individual PDF predictions for a specific product in the test period, to be compared to a uniform distribution.}
\end{center}
\end{figure}

\fig \ref{fig:invquant_mean} and \fig \ref{fig:invquant_dayofweek} show quantile profile plots for different variables on the x-axis, namely mean predictions and day of week, respectively. In the former case all product-location-day combinations in the test period are used to generate the statistics in the plot, while in the latter case all locations and days in the test period for item \texttt{FOODS\_3\_586} are aggregated. It can be seen that the different weekdays show different patterns and that there are slightly larger deviations for higher mean predictions (with fewer samples though).

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_yhat_mean}
\caption{\label{fig:invquant_mean} Quantile profile plot for mean predictions on the x-axis aggregated over all product-location-day combinations in the test period.}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_dayofweek}
\caption{\label{fig:invquant_dayofweek} Quantile profile plot for the different days of week on the x-axis aggregated over all locations and days in the test period for a specific product.}
\end{center}
\end{figure}

The quantitative results for the CDF accuracies of our PDF predictions for the different metrics described in sec. \ref{sec:cdf_acc}, calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf:acc}. As benchmark, we compare these to the results of a Poisson model assumption needing only the mean as single parameter. Our negative binomial PDF predictions show a significant improvement over the simpler Poisson model, which uses the same mean predictions, for each of the metrics.

\begin{table}[h!]
\begin{center}
\caption{CDF accuracies for negative binomial and Poisson PDF predictions for different metrics calculated over all product-location-day combinations in the test period.}
\label{tab:cdf:acc}
\begin{tabular}{c|c|c}
 & \textbf{NBD} & \textbf{Poisson} \\
\hline
\textbf{EMD} & 0.65 & 0.71 \\
\textbf{KL\_2} & 0.62 & 0.79 \\
\textbf{KL\_e} & 0.74 & 0.86 \\
\textbf{JSD\_2} & 0.92 & 0.94 \\
\textbf{JSD\_e} & 0.95 & 0.96
\end{tabular}
\end{center}
\end{table}
